The unprecedented energies and luminosities provided by the LHC have allowed testing our physical theories in never-before-seen regimes and with remarkable precision.
With data from collisions produced by the LHC gathered with the ATLAS detector\footnote{And CMS.}, the final piece of the Standard Model, the Higgs boson, has been confirmed.
The Standard Model has therefore withstood all experimental tests of its predictions, and there are no remaining undiscovered particles or forces predicted by the SM.

However, there are good motivations for believing the SM does not tell the full story, and it is almost certain that the LHC will play a major role in uncovering the next chapter.
It is therefore vitally important to put the data provided by the LHC to as best use as possible, in order to advance our understanding of nature and lead to the next big discovery.

The use of jets in novel and interesting ways is one crucial way of making best use of LHC data - it is again almost certain that the next big discovery will involve jets in some way.

The original research efforts described in this Thesis contribute to providing means by which this endeavor can be fulfilled.
Chapter~\ref{ch:HBSM} describes a search with multiple jets in the final state for new particles via their interactions with the Higgs boson.
Tests of the precise properties of the Higgs boson like these present a major remaining test for the SM.
The search in Chapter~\ref{ch:CWoLa} uses the underlying features of jets in a novel way to perform a generic search for new resonant particles.
These innovations in analysis techniques will enhance the potential of the LHC to be used to discover new unexpected physics phenomena.
Finally, Chapters~\ref{ch:NI} and~\ref{ch:GenNI} present new understandings and methods in the reconstruction of jets.
These improvements in the performance of jets enable their best use in measurements and searches with LHC data.

The search for new physics using jets continues, and the progress made in this Thesis will contribute to these efforts now and in the future.
